### 2024-12-09 draft.Rmd
Update draft.Rmd

---
title: " Group Report"
output:
  pdf_document: default
  html_notebook: default
editor_options: 
  markdown: 
    wrap: 72
---

Members: Tracy Gan, Zihua Li, Xinya Lu, Matilda Lucas

## I. Introduction

Education serves as a cornerstone for societal development, with
academic performance often reflecting both individual and systemic
success. This study explores the extent to which social and educational
factors influence student achievement, drawing on data from two
Portuguese secondary schools during the 2005--2006 academic year. The
dataset incorporates diverse variables, including parental education
levels, aspirations for higher education, and absences, alongside
demographic and school-related factors. By examining these elements, the
research aims to uncover key contributors to academic performance and
offer actionable insights for improving educational outcomes (Cortez &
Silva, 2008).

The broader context highlights the persistent challenges faced by
students in lower socioeconomic households, where access to resources
and parental involvement often differ significantly from their higher
socioeconomic status counterparts. For instance, previous studies
underscore the importance of early educational exposure, the quality of
home learning environments, and access to extracurricular activities as
critical components for academic success (Milne, 2005). Factors such as
parental education, the presence of role models, and emotional support
further differentiate students' abilities to thrive academically, even
within economically constrained households (Milne, 2005).

Motivated by these findings, this report investigates whether specific
social and educational variables significantly influence academic
performance. Using statistical techniques, including hypothesis testing
and logistic regression, the study seeks to quantify these relationships
and evaluate their impact. Ultimately, this research contributes to a
deeper understanding of how systemic and familial dynamics shape
educational achievement, with the aim of identifying interventions that
can enhance equity and access in education.

## II. Analysis

### A. Exploratory data analysis

```{r include = FALSE}
library(car) 
library(ggplot2) 
library(reshape2) 
library(tidyverse) 
library(broom) 
library(caret) 
library(RColorBrewer) 
library(leaps) 
library(rmarkdown)
library(knitr)
library(evaluate)

```

In our analysis, we aim to model student performance as the response
variable. The initial dataset includes performance scores from two
periods, G1 and G2. However, some students have a score of zero in one
of these periods, indicating they were absent. To ensure the accuracy of
our analysis, we created a new variable, G1_G2_Average, which represents
the average performance across both periods. We also removed data for
students who were absent in one period, as their absences should not
influence the measure of student performance.

```{r include= FALSE}
# load the dataset
data <- read.csv("dataset.csv")
head(data, 1)
cat("Table 1. Preview of the High School Alcoholism and Academic Performance dataset.\n")
```

```{r echo = FALSE, fig.cap = "Table 2. Preview of the cleaned High School Alcoholism and Academic Performance dataset.\n"}
# clean the dataset, and store the clean dataset
missing_summary <- sapply(data, function(x) sum(is.na(x)))
data[is.na(data)] <- lapply(data[is.na(data)], function(x) ifelse(is.numeric(x), median(x, na.rm = TRUE), "unknown"))

# Add a new column 'G1_G2_Average' that is the average of G1 and G2
data$G1_G2_Average <- rowMeans(data[c("G1", "G2")], na.rm = TRUE)

# Remove rows where either G1 or G2 is equal to 0
data <- subset(data, G1 != 0 & G2 != 0)
head(data,1)
cat("Table 1. Preview of the cleaned High School Alcoholism and Academic Performance dataset.\n")
```

```{r echo = FALSE, fig.cap = "  Distribution of average of G1 and G2.\n", fig.width=4, fig.height=2}
# Distribution histogram of the Average of G1 and G2
hist <- ggplot(data, aes(x = G1_G2_Average)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of average of G1 and G2", x = "Average", y = "Count") +
  theme_minimal()
hist
#cat("Figure 1. Distribution of average of G1 and G2.\n")
```

Figure 1 illustrates how the average grades are continuously
distributed. We can see most grades concentrated around the center and
there is no outliers or invalid data after we clean our dataset.

```{r include = FALSE}
full_model <- lm(G1_G2_Average ~ .-G1-G2, data = data)
summary(full_model)
```

```{r echo = FALSE, fig.cap="\n. Correlation heatmap.\n", fig.width=4, fig.height=4 }
#setup the covariance matrix
cor_matrix <- cor(data[, sapply(data, is.numeric)], use = "complete.obs")
#print(cor_matrix)
cor_data <- melt(cor_matrix)
ggplot(cor_data, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation Heatmap", x = "", y = "")
#cat("\nFigure 2. Correlation heatmap.\n")
```

Figure 2 uses a color gradient to represent the strength and direction
of correlations. We can see Medu, Fedu, and studytime might has a
stronger correlation with G1_G2_Average.

```{r echo = FALSE, fig.cap = "\n. Residual plot for full model.\n", fig.width=5, fig.height=4}
# residual plot again fitted values for full model
plot(full_model$fitted.values, residuals(full_model),
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 3. Residual plot for full model.\n")
```

Based on Figure 3, the assumption of constant variance might be
violated, as the plot appears to have a funnel-shaped pattern.

```{r echo = FALSE, fig.cap = "\n. Q-Q plot for full model.\n", fig.width=5, fig.height=4}
# QQ-plot for full model
qqnorm(residuals(full_model))
qqline(residuals(full_model), col = "red")
#cat("\nFigure 4. Q-Q plot for full model.\n")
```

Based on Figure 4, the residuals of the full model appear to follow a
normal distribution.

```{r include = FALSE}
# get the VIF values
vif_values <- vif(full_model)
print(vif_values)
```

VIF quantifies how much multicollinearity inflates the variance of a
regression coefficient. From the calculated VIF values, no variable
exceeds a value of 5, suggesting that multicollinearity is unlikely to
be a concern in the model.

```{r include = FALSE}
#split data
set.seed(11)
data <- 
    data %>% 
    mutate(id = row_number())

training_data <- 
    data %>% 
    slice_sample(prop = 0.70, replace = FALSE)

testing_data <- 
    data %>%
    anti_join(training_data, by = "id") %>%
    select(-"id")

training_data <- 
    training_data %>% 
    select(-"id")
```

To select the most appropriate covariates for our model, we will perform
backward selection. First, we will split the data into a training set
and a testing set. The training set will be used for the backward
selection process, while the testing set will be used to build and
evaluate the final model.

```{r echo = FALSE}
#backward selection to choose the covariates
your_backward_sel <- regsubsets(
  x = G1_G2_Average ~ .-G1-G2,  
  nvmax = 4,       
  data = training_data,         
  method = "backward"          
)

#your_backward_sel

your_bwd_summary <- summary(your_backward_sel)

your_bwd_summary_df <- tibble(
  n_input_variables = 1:4,          
  R2 = your_bwd_summary$rsq,
  RSS = your_bwd_summary$rss,
  ADJ.R2 = your_bwd_summary$adjr2,
  BIC = your_bwd_summary$bic,
  Cp = your_bwd_summary$cp
)

your_bwd_summary_df
#your_bwd_summary
```

Selecting a maximum of 4 variables ensures that the final model remains
interpretable and avoids unnecessary complexity, making it easier to
explain and derive actionable insights. Additionally, limiting the
number of predictors helps reduce the risk of overfitting.

Based on the result of backward selection, we will use student's school
(school), mother's education level (Medu) , individual desire to purse a
degree or not (higher), and number of school absences (absence) as our
explantory variables

```{r include = FALSE}
# linear model with no transformation
model <- lm(G1_G2_Average ~ school + Medu + higher + absences, data = testing_data)
summary(model)
```

Then we create a linear model between the avarage student performance
and the four covariates that we choose.

```{r echo = FALSE, fig.cap = "\n. Q-Q plot for actual model.\n", fig.width=5, fig.height=4}
qqnorm(residuals(model))
qqline(residuals(model), col = "red")
#cat("\nFigure 5. Q-Q plot for actual model.\n")
```

Based on Figure 5, the residuals of our actual model appear to generally
follow a normal distribution, although the tail on right is slightly
off.

```{r echo = FALSE, fig.cap ="\n. Residual plot for actual model.\n", fig.width=5, fig.height=4}
plot(model$fitted.values, residuals(model), 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 6. Residual plot for actual model.\n")
```

Based on Figure 6, the residual plot for the fitted model indicates that
the constant variance assumption is likely violated. To address this
issue, we decided to apply a log transformation to the fitted values.

```{r include = FALSE}
# linear model with log transformation
model2 <- lm(log(G1_G2_Average) ~ school + Medu + higher + absences, data = testing_data)
summary(model)
```

```{r echo = FALSE, fig.cap ="\n. Q-Q plot for log-transformed model.\n", fig.width=5, fig.height=4 }
qqnorm(residuals(model2))
qqline(residuals(model2), col = "red")
#cat("\nFigure 7. Q-Q plot for log-transformed model.\n")
```

```{r echo = FALSE, fig.cap ="\n. Residual plot for log-transformed model.\n", fig.width=5, fig.height=4 }
plot(model2$fitted.values, residuals(model2), 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 8. Residual plot for log-transformed model.\n")
```

After applying the log transformation to the model, the model fit
improved significantly. The new residual plot shows that the residuals
are more randomly scattered around the zero line, indicating a better
fit.

## III. Model Inference

In order to investigate the relationship between student performance scores (G1_G2_Average) and the variables of interest, hypothesis testing was conducted using logistic regression models and confidence intervals were evaluated. The objective of this study was to analyze the effect of the following variables on student performance:

- School 
- Medu
- Higher 
- Absences

#### Hypothesis Testing

For each variable, we set the following hypotheses:
$$
\begin{aligned}
H_0 &: \beta_j = 0 \\
H_a &: \beta_j \neq 0
\end{aligned}
$$

where $j = 1, 2, 3, 4$.

The equation of model is given by
$$
y = \beta_0 + \beta_1 X_{schoolMS} + \beta_2 X_{Medu} + \beta_3 X_{higheryes} + \beta_4 X_{absences}
$$

The Wald's statistic $z_j$ is given by the formula:

$$
z_j = \frac{\hat{\beta}_j}{SE(\hat{\beta}_j)}
$$

Here, we perform hypothesis testing and interpret the p-values for each predictor variable (School, Medu, Higher, and Absences) to evaluate their significance in association with the student performance scores. The code is adapted to include confidence intervals for a more comprehensive understanding of the results.

```{r echo = FALSE}
results <- model %>%
  tidy(conf.int = TRUE)
print(results)
```

### Results
#### Coefficient Significance $(α = 0.05)$ / Confidence Intervals $(95\%)$
- **Intercept**: $\beta_0\$ \( [8.97, 11.4] \)

   We reject $H_0$ since the $p-value$ is $3.41 \times 10^{-38}$. This indicates that the intercept is statistically significant, suggesting that the baseline levels of the predictors have a meaningful contribution to the average student performance.

   Besides that we are 95% confident that the true value of the intercept lies within $[8.97, 11.4]$. Since the range is entirely > 1, it also suggests that the baseline group is associated with higher levels of student performance compared to other predictor categories.

- **School**: $\beta_1\$ \( [-2.58, -1.15] \)

  The $p-value$ $6.40 \times 10^{-7}$ is less than $0.05$, so we reject $H_0$. This result indicates that the type of school is significantly associated with average student performance.

  The confidence interval for School lies entirely below 0, which shows the same result.

- **Medu**: $\beta_2\$ \( [0.0418, 0.632] \)

  The $p-value$ $2.55 \times 10^{-2}$ is less than $0.05$, so we reject $H_0$. There is enough evidence to conclude that mother's education level is significantly associated with average student performance.

  The confidence interval didn't includes 0, also indicates evidence to conclude a relationship between mother’s education level (Medu) and higher student performance.

- **Higher**: $\beta_3\$ \( [0.826, 2.93] \)

  The $p-value$ $5.42 \times 10^{-4}$ is less than $0.05$, so we reject $H_0$. This suggests that planning to pursue higher education is significantly associated with improved average student performance.

  We can see the range $[0.826, 2.93]$ lies entirely > 0, also suggests a significant positive relationship between planning to pursue higher education degree and better student performance.

- **Absences**: $\beta_4\$ \( [-0.152, -0.0152] \)

  The $p-value$ $1.68 \times 10^{-2}$ is less than $0.05$, so we reject $H_0$. This indicates that absences are significantly associated with lower average student performance.

  The confidence interval lies entirely below 0, also suggests a significant negative association between the number of absences and higher student performance.

## IV. Conclusion

The objective of this analysis was to build an inferential model to
investigate whether social and educational factors influence academic
performance, specifically the variable G1_G2_Average which represents
the average academic performance of a student across both periods. Our
results from performing backward selection suggests that the student's
school (school), mother's education level (Medu), individual desire to
purse a degree or not (higher), and number of school absences (absence)
were important in influencing the students' academic performance.

### Model fit

From the linear model fitted with these variables, we can see the plot
of these models in Figure 6 showed us that our model is not an
appropriate fit. The plot showed presence of heteroscedasticity as there
is fan-shaped pattern, violating a key assumption of constant variance
of linear regression. Therefore we applied a log transformation of the
predictor variable, resulting in a better fitted model as seen in Figure
7 as the assumptions of linear regression are met.

From this model, we got an $R^2$ value of 0.259, meaning that our model 
explains 25.9% of the variance in log of (G1_G2_Average). Additionally, 
this $R^2$ value did not improve, and the $R^2$ value our model achieve 
is not that high indicating that the relationship between target and 
predictor might not have a log-linear relationship between them.

### Statistical Significance of Predictors

The summary of our model shows that all four variables are statistically
significant at a critical value of 0.05, indicating to us that these
four variables have a meaningful relationship with G1_G2_Average.
Specifically, \* schoolMS has a negative correlation with
log(G1_G2_Average) with estimate -1.42. This implies that being in
school MS (Mousinho da Silveira) as compared to school GP (Gabriel
Pereira) is associated with a decrease in log(G1_G2_Average) by 1.42
when other variables are constant \* Medu has a positive correlation of
0.215, implying that for a unit increase in Mother's educational level,
the log(G1_G2_Average) increases by 0.215 units. \* higherYes has a
positive correlation of 2.41, implying that an increase in 2.41 units in
log(G1_G2_Average) is observed for students planning to pursue a degree
compared to those who are not interested. \* absences has a negative
correlation of 0.131, implying that every additional day of absences is
associated with a decrease in log(G1_G2_Average) by 0.131 units.

### Implications

The findings from above suggests that higher maternal education and
desire to pursue a degree are associated with a higher G1_G2_Average,
meaning higher academic average performance of students, whilst
increasing number of absences are associated with lower academic average
of students. As school type also has a negative impact on the academic
average, this points to a difference in outcomes of students in these
different schools.

### Final Thoughts

In conclusion, although our model provides statistical significant
results for all four predictors, the low $R^2$ value of 0.259 suggests
that there might be other predictor variables influencing
log(G1_G2_Average) that are not captured in our analysis. Further
exploration and analysis of the predictor variables in the future may
help with defining these relationships.

## IV. Appendix
### References

Cortez, P., & Silva, A. M. G. (2008, April). Using data mining to
predict secondary school student performance.
<https://repositorium.sdum.uminho.pt/handle/1822/8024>

Milne, A. (2005). Factors of a Low-SES Household: What Aids Academic
Achievement. All Graduate Projects.
<https://digitalcommons.cwu.edu/graduate_projects/878>